---
title: "CHASE Test"
output: github_document
---
## Testing the CHASE hazardous substances assessment tool

This is a test of the CHASE tool, using two input datasets: one for level 3 assessment units and one for level 4 assessment units. Both datasets are based on the same indicator results but are aggregated at different spatial scales. A description of the preparation of these test data sets can be seen in [test_dataset.md](test_dataset.md).

The results of the test are stored in the [output](output/) folder.

The code contained in this markdown document is also available as a standalone R script in [src/CHASE_test.R](src/CHASE_test.R).
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path = "markdown_images/")
```

Load required functions
```{r load functions, message=FALSE}
require(tidyverse)
source('src/CHASE.R')
source('src/CHASE_functions.R')
```

Load indicator data for level 3 and level 4 assessments.
```{r load data, message=FALSE}
# load data for level 3 indicators
file3 <- "./input/assessmentdata_L3.csv"
df3 <- read.table(file3,sep=";",header=T)

# load data for level 4 indicators
file4 <- "./input/assessmentdata_L4.csv"
df4 <- read.table(file4,sep=";",header=T)

# show the head of df3
head(df3)
```

Do the assessments
```{r assessments, message=FALSE}
# Get assessment results for each assessment level
CHASE3<-CHASEassessment(df3) 
CHASE4<-CHASEassessment(df4)
```

The fourth item in each result list gives the overall results for each assessment unit.

```{r confidence results, message=FALSE}
resL3<-CHASE3[[4]]
resL4<-CHASE4[[4]]

# show the head of resL3
head(resL3)
```

Plot the variations in confidence for the overall assessment

```{r plot confidence results, message=FALSE}

p3<-ggplot(resL3) +
  geom_histogram(aes(x=ConfScore),binwidth=0.02) +
  ggtitle("Level 3") +
  theme_minimal()
plot(p3)

p4<-ggplot(resL4) +
  geom_histogram(aes(x=ConfScore),binwidth=0.02) +
  ggtitle("Level ") +
  theme_minimal()
plot(p4)
```

Save the results and the plots

```{r save results, message=FALSE}
ggsave(p3,file="output/histogram_confidence_L3.png",dpi=300,units="cm",height=15,width=15)
ggsave(p4,file="output/histogram_confidence_L4.png",dpi=300,units="cm",height=15,width=15)

save(CHASE3,file="output/test_results_L3.Rda")
save(CHASE4,file="output/test_results_L4.Rda")

```